Testing for the prevalence or transience of a shock effect in the post-shock setting

### 2021-07-01 prevalence_testing/postshocktest2021draft.tex
reorganized repository

\documentclass[11pt,3p,review,authoryear]{elsarticle}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amsfonts}
%\usepackage{amscd}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{mathtools}
\usepackage{natbib}
\usepackage{enumitem}
\usepackage{bm}
\usepackage[usenames]{color}
\usepackage{caption}
\usepackage{float}
\usepackage[caption = false]{subfig}
\usepackage{tikz}
\usepackage{multirow}
\usepackage[linesnumbered, ruled,vlined]{algorithm2e}
\usepackage{pdflscape}
\PassOptionsToPackage{hyphens}{url}
\usepackage{hyperref}

@@

\journal{International Journal of Forecasting}

\bibliographystyle{model5-names}
\biboptions{longnamesfirst}

@@

\begin{frontmatter}

\title{Minimizing post-shock forecasting error through aggregation of outside information}

\begin{abstract}
    We develop a forecasting methodology for providing credible forecasts for time series that have recently undergone a shock. We achieve this by borrowing knowledge from other time series that have undergone similar shocks for which post-shock outcomes are observed. Three shock effect estimators are motivated with the aim of minimizing average forecast risk. We propose risk-reduction propositions that provide conditions that establish when our methodology works. Bootstrap and leave-one-out cross validation procedures are provided to prospectively assess the performance of our methodology. Several simulated data examples, and two real data examples of forecasting Conoco Phillips and Apple stock price are provided for verification and illustration.
\end{abstract}

\begin{keyword}
  Data Integration \sep 
 Prospective forecasting \sep 
 Risk reduction \sep 
Residual bootstrap \sep 
Cross validation 
\end{keyword}

\end{frontmatter}

\section{Introduction}

We provide forecasting adjustment techniques with the goal of lowering overall forecast error when the time series under study has undergone a structural shock. We focus on the specific setting in which a structural shock has occurred and one desires a prediction for the post-shock response at the next time point. Standard forecasting methods may not yield accurate predictions in the presence of such structural shocks \citep{baumeister2014real}. This is a general problem that has many real life applications. For example, one may be interested in forecasting the stock price of a company tomorrow after hearing terrible or great news about the company after trading has concluded for the day. Companies may be interested in forecasting the demand of their products to adjust production after they were involved in a brand crisis, but they only have recent sales data for which the company is operating well. All is not lost in this forecasting setting, one may be able to supplement the present forecast with past data borrowed from other time series which contain similar structural shocks. The core idea of our methodology is to sensibly aggregate similar past realized shock effects which arose from other time series, and then incorporate the aggregated shock effect estimator into the present forecast. Our method of combining shock effects embraces ideas from conditional forecasting \citep{baumeister2014real, kilian2017structural}, time series pooling using cross-sectional panel data \citep{ramaswamy1993empirical, pesaran1999pooled, hoogstrate2000pooling, baltagi2008forecasting, koop2012forecasting, liu2020forecasting}, forecasting with judgement and models \citep{svensson2005monetary, monti2008forecast}, synthetic control methodology \citep{abadie2010synthetic, agarwal2020two}, and expectation shocks \citep{croushore2006data, baumeister2014general, clements2019measuring}.

We study the post-shock forecasting problem in the context of additive shock effects in linear autoregressive models. In this post-shock forecasting setting, the researcher has a time series of interest which is known to have recently undergone a structural shock, and the post-shock response is not observed. In this setting, we treat the additive shock effect as an observed random effect, obtained from a distribution of similar shocks, that is parameterized as a conditional fixed effect within an autoregressive model. The shock effect is then estimated using ordinary least squares (OLS). The researcher must move beyond the modeling paradigm that they were previously working under to accommodate this new shock effect \citep{monti2008forecast, svensson2005monetary}. One method for estimating the shock effect is to produce a conditional point forecast where a sequence of non-zero future structural shocks are conditioned upon and estimated \citep{baumeister2014real}. Such conditional point forecasts are appropriate when the shock sequence considered is within the range of historical experience \citep{kilian2017structural}. On the other hand, our methodology allows for the inclusion of outside data sources and covariates into this conditional forecasting framework, provided that the shock effects from outside data sources arose from the same data generating process as the shock effect under study. For further differences of assumptions on shocks, our methodology allows for unprecedented shocks and no observation of shocks from the recent past. Our shock effect setting is also similar to the data generating mechanism that  \cite{tsay1986time} studied for ARMA models. \cite{tsay1986time} studied additive outliers that contaminates the data at only one or multiple specific points, and innovational outliers that impact the subsequent responses. Our shock effect of interest is similar to innovational  outliers  in the sense that it is an additive effect in the model of response only at a particular time point but it impacts the subsequent responses through an autoregressive structure.

In our methodological framework, the researcher first creates a donor pool of time series which have undergone similar structural shocks in the past. Similarity in this context refers to the shock effects as arising from a distribution that has existing second moments, with extensions to variances and means that are linear functions of common covariates. The unobserved shock effect in the time series of interest is assumed to arise from the same distribution. Construction of the donor pool and the desired aims of the donor pool has precedent in synthetic control methodology (SCM) \citep{abadie2010synthetic}. As in SCM, care is needed when forming the donor pool of time series. However, there are several key differences between our framework and SCM. Unlike in SCM, we are not trying to estimate the causal effect of an intervention. Also, we assume that the time series forming the donor pool are independent from the time series under study before the timing of the shock. This rules out panel data applications where the time series of interest and those forming the donor pool evolve in tandem. When there are large time differences between the time series under study and those forming the donor pool, then the dependence will be negligible in practice. This is the case in the Conoco Phillips and Apple stock forecasting examples.

We then estimate the shock effects that are present in the time series forming the donor pool for which the post-shock responses are observed. These estimated shock effects are then aggregated. Finally, this aggregated estimator is added to the forecast for the yet to be realized post-shock response of the time series of interest. Our adjustment is similar to various adjustment methodologies suggested by \cite{clements1996intercept} that use the residual at end of observed time horizon for $h$-ahead forecasts. However, the approach of \cite{clements1996intercept} is not robust to unprecedented and unobserved shock since the shock signal is not embedded in the data. Under a similar setting that studies equilibrium shifts, \cite{castle2015robust} propose robust general approaches for forecasting in vector equilibrium-correction models. However, their forecasting methodology requires conditioning on the past data that contain shock signals. This approach is therefore not applicable in the post-shock setting. The shock effects in our post-shock forecasting framework are similar to ``expectation shocks,'' studied in \cite{clements2019measuring}. \cite{clements2019measuring} allows for consistent estimation of expectation shocks under a vector autoregressive model, possibly involving an instrumental variable approach, as in \cite{croushore2006data}. In our context, the yet to be observed shock effect of interest arises as a random effect whose assumed generative distribution can only be partially estimated. This partial estimation only uses shock effect estimates from the time series forming the donor pool.

In this article, we will assume a simple autoregressive data generating process  with a general random-effect structure. These specifications are similar to the ``$K$ latent pooling'' framework of \cite{ramaswamy1993empirical}. However, our model formulation is more general than \cite{ramaswamy1993empirical}. In our model, the donor pool can consist of dependent time series, but the time series forming the donor pool are required to be independent from the time series of interest. However, mutual independence among time series in the donor pool can aid prospective evaluation of the reliability of our method. We consider three aggregation techniques: simple averaging, inverse-variance weighted averaging, and similarity weighting. The latter technique is similar to the weighting in SCM \citep{abadie2010synthetic}. The considered adjustment strategies all target the mean of the shock effect distribution. Such an estimation strategy can reduce mean squared error (MSE) when variation in the shock effect distribution is small relative to the mean. We provide risk-reduction propositions that detail the conditions when the adjusted forecasts will work better than the original forecast. The involved parameters in the risk-reduction propositions can be estimated by a residual bootstrap procedure that we develop. We also motivate a simple leave-one-out cross validation procedure which can prospectively assess the performance of our shock effect adjustment estimators. This prospective assessment does not require the observation of the post-shock response. Our Monte Carlo simulation results show that the risk-reduction propositions are nearly perfectly correct when the model for the shock effects is identified well with appropriate covariates under a fixed design.  We demonstrate the utility of our methodology with two real data analyses. The first of these focuses on forecasting the stock price of Conoco Phillips after March 9th, 2020, in the midst of the COVID-19 pandemic and an oil supply war. The second of these focuses on forecasting the stock price of Apple Inc. after the company releases their latest installment of the MacBook on November 10th, 2020. We will show that our proposed adjustment estimators yield much better results than no adjustment in these examples. The Conoco Phillips example provides an empirical application of our methods when the shock is anticipated yet unprecedented in nature, arising from a global pandemic and an oil supply shock. The Apple example provides a case for post-shock prediction when the shock has precedent, we forecast the stock price after the latest release of the MacBook. We now elaborate our framework for post-shock prediction, we start with an illustration.

\begin{figure*}[t!]
  \begin{center}
    \subfloat[]{\includegraphics[height = 5.5cm]{FIG1a.pdf}}\\
     \subfloat[]{\includegraphics[height = 5.5cm]{FIG1b.pdf}}
  \end{center}
     \caption{The time series experience a shock at $T_1^*+1=126$ with true shock effect $\alpha = 9.21$. (a) presents the comparison of the prediction without adjustment and one that uses simple averaging of estimated shock effects given a donor pool of $n=40$ time series. (b) shows the histogram for the least square estimates $\hat{\alpha}_i$s used in estimating shock $\alpha_1$ for $i = 2, \ldots, 41$.  The magenta dots represent $\hat{\alpha}_i$ from the time series forming the donor pool. $\hat{y}^{1}_{T_1^*+1}$ is the prediction without any adjustment, and $\hat{y}^{2}_{T_1^*+1}$ is the forecast with additive adjustment of $\hat{\alpha} = n^{-1}\sum_{i=2}^{n+1}\hat\alpha_i$. In this example, theses forecasts differ by $\hat{\alpha}=10.22$. It is clear that $\hat{y}^{2}_{T_1^*+1}$ performs better than $\hat{y}^{1}_{T_1^*+1}$.}\label{figure1}
\end{figure*}

\section{Setting}
\label{setting}

We will suppose that a researcher has time series data ($y_{i,t}$,$\x_{i,t}$), for $t = 1, \ldots,  T_i$ and $i = 1, \ldots, n+1$, where $y_{i,t}$ is a scalar response and $\x_{i,t}$ is a vector of covariates that are revealed to the analyst prior to the observation of $y_{1,t}$.  Suppose that the analyst is interested in forecasting $y_{1,t}$, the first time series in the collection. We will suppose that specific interest is in forecasting the response after the occurrence of a structural shock. To gauge the performance of forecasts, we consider forecast risk in the form of MSE,
$$
  R_T = \frac{1}{T}\sum_{t=1}^T\E{\hat y_{1,t} - y_{1,t}}^2,
$$
and root mean squared error (RMSE), given by $\sqrt{R_T}$, in our analyses. In this article, we focus on post-shock prediction where forecasting methods only differ at the next future time point. Thus the MSE reduces to the magnitude $\E{\hat y_{1,t} - y_{1,t}}^2$.

Our post-shock forecasting methodology will consist of selecting covariates $\x_{i,t}$, constructing a suitable donor pool of candidate time series that have undergone similar structural shocks to the time series under study, and specifying a model for the time series ($y_{i,t}$,$\x_{i,t}$), for $t = 1, \ldots,  T_i$ and $i = 1, \ldots, n+1$. In this article, we consider a  special case that the model includes random effects for the covariates, autoregressive structure, and an additional shock effect, whose presence or absence is given by the indicator variable $D_{i,t}$.  We will assume that the donor pool time series are independent of the time series under study. The details of this model are in the next section.

Figure \ref{figure1} provides a simple intuition of the practical usefulness of our proposed methodology. This figure depicts a time series that experienced a shock at time point $T_1^*+1 = 126$. It is supposed that the researcher does not have any information after $T_1^*$, but does have observations from a donor pool of forty time series that have previously undergone similar shocks for which post-shock responses are recorded. Similarity in this context means that the shock effects are random realizations from a common distribution. In this example, the mean of the estimated shock effects is taken as a shock effect estimator for the time series under study. Forecasts are then made by adding this shock effect estimator to the estimated response values obtained from the estimation procedure that ignores the shock. It is apparent from Figure \ref{figure1} that adjusting forecasts in this manner 1) leads to a reduction in forecasting risk; 2) does not fully recover the true shock effect. We evaluate the performance of this post-shock forecasting methodology throughout this article; we outline situations for when it is expected to work and when it is not.

@@

We now describe the assumed autoregressive models with random effects for which post-shock aggregated estimators are provided. The basic structures of these models are the same for all time-series in the analysis, the differences between them lie in the setup of the shock effect distribution.

Let $I(\cdot)$ be an indicator function, $T_i$ be the time length of the time series $i$ for $i = 1, \ldots, n+1$, and $T_i^*$ be the time point just before the one when the shock is known to occur, with $T_i^* < T_i$.  For $t= 1, \ldots, T_i$ and $i = 1, \ldots, n+1$, the model $\mc{M}_1$ is defined as
\begin{align}
\mc{M}_1 \colon y_{i,t} =\eta_i +\alpha_i D_{i,t} + \phi_i y_{i, t-1} + \theta_i'\mbf{x}_{i,t} + \varepsilon_{i,t}\label{equation1}
\end{align}
 where $D_{i,t} = I(t = T_i^* + 1)$ 
and $\x_{i,t} \in \R^{p}$ with $p \geq 1$.  We assume that the 
$\mbf{x}_{i,t}$s are fixed. Let $|x|$ denote the absolute value of $x$ for $x\in \reals$. For $i = 1, \ldots, n+1$ and $t=1, \ldots, T_i$, the random effects structure for $\mc{M}_1$ is:
\begin{align*}
  \eta_i &\simiid \mc{F}_{\eta} \text{ with }  \; \mrm{E}_{\mc{F}_{\eta}}(\eta_i) = 0, \mrm{Var}_{\mc{F}_{\eta}}(\eta_i)  = \sigma^2_{\eta}\\
  \phi_i &\simiid \mc{F}_{\phi} \text{ where } |\mc{F}_{\phi}| < 1, \\
   \theta_i &\simiid \mc{F}_{\theta} \text{ with }  \; \mrm{E}_{\mc{F}_{\theta}}(\theta_i) = \mu_{\theta}, \mrm{Var}_{\mc{F}_{\theta}}(\theta_i)  = \Sigma^2_{\theta} \\
\alpha_i &\simiid \mc{F}_{\alpha} \text{ with }  \; \mrm{E}_{\mc{F}_{\alpha}}(\alpha_i) = \mu_{\alpha}, \mrm{Var}_{\mc{F}_{\alpha}}(\alpha_i)  = \sigma^2_{\alpha}  \\
\varepsilon_{i,t} & \simiid  \mc{F}_{\varepsilon_i} \text{ with }  \; \mrm{E}_{\mc{F}_{\varepsilon_i}}(\varepsilon_{i,t}) = 0, \mrm{Var}_{\mc{F}_{\varepsilon_i}}(\varepsilon_{i,t})  = \sigma^2_i  ,  \\
\eta_i &\indep  \alpha_i \indep \phi_i \indep \theta_i \indep \varepsilon_{i,t},
\end{align*}
where $\indep $ denotes the independence operator. Notice that $\mc{M}_1$ assumes that $\alpha_i$ are iid with $\E{\alpha_i}=\mu_{\alpha}$ 
for $i = 1, \ldots, n+1$. We also consider a model where the shock effects are linear functions of covariates with an additional additive mean-zero error. For $i = 1, \ldots, n+1$, the random effects structure for this model (model $\mc{M}_2$) is:
\begin{align}
\mc{M}_2 \colon \begin{array}{l}
  y_{i,t} =\eta_i +\alpha_i D_{i,t} + \phi_i y_{i, t-1} + \theta_i'\mbf{x}_{i,t} + \varepsilon_{i,t}\\[.2cm]
  \; \alpha_i = \mu_{\alpha}+\delta_{i}'\mbf{x}_{i, T_i^*+1}+ \t{\varepsilon}_{i},
\end{array}\label{model2}
\end{align}
 where the added random effects are
\begin{align*}
\t{\varepsilon}_{i} &\simiid  \mc{F}_{\t{\varepsilon}} \text{ with }\mrm{E}_{\mc{F}_{\t{\varepsilon}}}(\t{\varepsilon}_{i})=0, \mrm{Var}_{\mc{F}_{\t{\varepsilon}}}(\t{\varepsilon}_{i})=\sigma^2_{\alpha}\\
\eta_i &\indep  \alpha_i \indep \phi_i \indep \theta_i \indep \varepsilon_{i,t} \indep \t{\varepsilon}_{i}.
\end{align*} 
We further define 
$\tilde{\alpha}_i=\mu_{\alpha}+\delta_i'\mbf{x}_{i, T_i^*+1}$. 
We will investigate the post-shock aggregated estimators in $\mc{M}_2$ 
in settings where $\delta_i$ is either fixed or random. 
We let $\mc{M}_{21}$ denote model $\mc{M}_{2}$ with $\delta_i = \delta$ for $i= 1, \ldots, n+1$, 
where $\delta$ is a  fixed unknown parameter.
We let $\mc{M}_{22}$ denote model $\mc{M}_{2}$ with the following random effects :
\begin{align*}
&\delta_i\simiid  \mc{F}_{\delta} \text{ with }\mrm{E}_{\mc{F}_{\delta}}(\delta_i)=\mu_{\delta}, \mrm{Var}_{\mc{F}_{\delta}}(\delta_i)=\Sigma_\delta \\
   & \delta_i  \indep \t{\varepsilon}_{i}.
\end{align*}
We further define the parameter sets
\begin{align}
\begin{array}{l}
 \;\, \Theta = \{(\eta_i, \phi_i, \theta_i, \alpha_i, \mbf{x}_{i,t}, y_{i,t-1}, \delta_i)\colon    t= 1, \ldots, T_i, i = 2, \ldots, n +1\} \\
   \Theta_1 = \{(\eta_i, \phi_i, \theta_i, \alpha_i, \mbf{x}_{i,t}, y_{i,t-1}, \delta_i)\colon  t= 1, \ldots, T_i, i = 1\}
\end{array}\label{parameter} 
\end{align}
where $\Theta$ and $\Theta_1$ can adapt to $\mc{M}_1$ by dropping $\delta_i$. We assume this for notational simplicity.

\subsection{Forecast}
\label{forecast}
In this section we show how post-shock aggregate estimators improve upon standard 
forecasts that do not account for the shock effect.

More formally, we will consider the following candidate forecasts: 
\begin{align*}
  &\text{Forecast 1}: \hat y_{1,T_1^*+1}^1 = \hat\eta_1 
    + \hat\phi_1 y_{1,T_1^*} + \hat\theta_1'\x_{1,T_1^*+1} 
    , \\
  &\text{Forecast 2}: \hat y_{1,T_1^*+1}^2 = \hat\eta_1 
    + \hat\phi_1 y_{1,T_1^*} + \hat\theta_1'\x_{1,T_1^*+1} 
    + \hat{\alpha},
\end{align*}
where $\hat\eta_1$, $\hat\phi_1$, and $\hat\theta_1$ are all OLS estimators of $\eta_1$, $\phi_1$, and $\theta_1$, respectively, and $\hat{\alpha}$ is some form of estimator for the shock effect of time series of interest, i.e., $\alpha_1$. 
The first forecast ignores the presence of $\alpha_1$ while the second forecast 
incorporates an estimate of $\alpha_1$ that is obtained from the other independent forecasts under study.

%Note that the two forecasts do not differ in their predictions for $y_{1,t}$, $t = 1,\ldots T_1^*$. Instead, they only differ in predicting $y_{1,T_1^*+1}$. 
Throughout the rest of this article we highlight when the information from the time series donor pool, indexed by $\{y_{i,t} \colon t = 1,\ldots,T_i, i = 2,\ldots,n+1\}$, can be used to construct a shock effect estimator $\hat\alpha$ in which Forecast 2 beats Forecast 1 and vice-versa. We will consider the different dynamic panel models $\mc{M}_1$, $\mc{M}_{21}$, and $\mc{M}_{22}$. Improvement will be measured by assessing the reduction in risk that Forecast 2 offers over Forecast 1. We outline the theoretical details of risk-reduction in Section \ref{properties}. We specifically focus on predictions for $y_{1,T_1^*+1}$, the first post-shock response. It is important to note that, in general, $\hat{\alpha}$ does not converge to $\alpha_1$ in any sense.  Despite this shortcoming, adjustment of the forecast for $y_{1,T_1^*+1}$ through the addition of $\hat{\alpha}$ has the potential to lower forecast risk under several conditions, and for several estimation techniques.

\subsection{Shock effect estimators}
\label{constructionofestimators}

We now construct the aggregated estimators of the shock effects that appear in Forecast 2 (see Section \ref{forecast}). We use these to forecast response values $y_{1, T_1^*+1}$ assuming that $T_1^*$ is known.  First, we introduce the procedures of parameter estimation for $\mc{M}_1$, $\mc{M}_{21}$, and $\mc{M}_{22}$ (see Section \ref{modelsetup}). For $i = 2, \ldots, n+1$, all parameters in this model will be estimated via ordinary least squares (OLS) using historical data indexed by $t = 1, \ldots, T_i$. For $i = 1$, we estimate all the parameters, excluding $\alpha_1$, using OLS for time periods indexed by $t=1, \ldots, T_1^*$. Let $\hat{\alpha}_i$ be the OLS estimate of $\alpha_i$, $i = 2, \ldots, n+1$.  Note that parameter estimation under $\mc{M}_1$ is the same as that for $\mc{M}_{21}$ or $\mc{M}_{22}$.  We emphasize that the $\alpha_i$s are realizations from a random effects distribution, and that the OLS estimators $\hat\alpha_i$ assumes that these realized values are conditional fixed effects in an autoregressive model, as in \eqref{equation1} and \eqref{model2}.

The candidate estimators for $\alpha_1$ are now introduced. Define the adjustment estimator for time series $i=1$ by 
\begin{equation} \label{adjusted}
  \hat{\alpha}_{\text{adj}} = \frac{1}{n}\sum_{i=2}^{n+1}\hat{\alpha}_i,
\end{equation}
We can use $\hat{\alpha}_{\text{adj}}$ as an estimator for 
the unknown $\alpha_1$ term for which no meaningful estimation information 
otherwise exists. It is intuitive that $\hat{\alpha}_{\rm adj}$ should perform 
well under $\mc{M}_1$ where we assume that $\alpha_i$s share the same mean 
for $i= 1, \ldots, n+1$. However, it can also be shown that 
$\hat{\alpha}_{\rm adj}$ may be less favorable in $\mc{M}_{21}$ 
and $\mc{M}_{22}$, which will be discussed in detail in Section \ref{properties}.

We also consider the inverse-variance weighted estimator 
in practical settings where the $T_i$s and $T_i^*$s vary greatly across $i=2, \ldots, n+1$. 
The inverse-variance weighted estimator is defined as 
\begin{align*}
  \hat{\alpha}_{\rm IVW} = \frac{\sum_{i=2}^{n+1} \hat{\alpha}_i / \hat{\sigma}_{i\alpha}^2}{\sum_{i=2}^{n+1} 1/\hat{\sigma}_{i\alpha}^2},
  \quad \text{ where } \quad  \hat{\sigma}_{i\alpha}^2 = \hat{\sigma}^2_i( \mathbf{M}_i'\mbf{M}_i)_{22}^{-1},
\end{align*}
where  $\hat{\alpha}_i$ is the OLS estimator of $\alpha_i$, $\hat{\sigma}_i$ is the residual standard error from OLS estimation, and $\mbf{M}_i$ is the design matrix for OLS with respect to time series for $i = 2, \ldots, n+1$. Note that since $\sigma$ is unknown, estimation is required and the numerator and denominator terms are dependent in general. However, $\hat{\alpha}_{\rm IVW}$ can be a reasonable estimator in practical settings, especially when the $T^*_i$s vary considerably. We do not provide closed form expressions for $\E{\hat{\alpha}_{\rm IVW}}$ and $\var{\hat{\alpha}_{\rm IVW}}$ but empirical performance of $\hat{\alpha}_{\rm IVW}$ is assessed via Monte Carlo simulation (see Section \ref{simulation}).

We now motivate a weighted-adjustment estimator appropriate for model $\mc{M}_{21}$ and $\mc{M}_{22}$. Our weighted-adjustment estimator is inspired by the weighting techniques in synthetic control methodology (SCM) developed in \cite{abadie2010synthetic}. However, our weighted-adjustment estimator is not a pure causal estimator, and our estimation premise is a reversal of the SCM context in \cite{abadie2010synthetic}. In our context, we predict the unobserved post-shock response $y_{1,T_1^*+1}$ (intervention is absent) using post-shock information from other time series whose post-shock responses are all observed (intervention is present).

We use similar notation as that in \cite{abadie2010synthetic} to describe our weighted-adjustment estimator. Consider a $\mbf{W} \in \R^n$ weight vector $\mbf{W}=(w_2, \ldots, w_{n+1})'$, where $w_i\in [0,1]$ for all $i = 2, \ldots, n+1$. Construct
\begin{align*}
 \mbf{X}_1 &= \mbf{x}_{1, T_1^*+1}',
  \quad  
  \mbf{X} = \begin{pmatrix}
    \mbf{x}_{2, T_2^*+1}' \\
    \vdots \\

>>>>

\documentclass[11pt]{article}

% use packages
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amsfonts}
%\usepackage{amscd}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{mathtools}
\usepackage{natbib}
\usepackage{enumitem}
\usepackage{url}
\usepackage{authblk}
\usepackage{bm}
\usepackage[usenames]{color}
\usepackage{hyperref}
\usepackage{geometry}
\usepackage{caption}
\usepackage{float}
\usepackage[caption = false]{subfig}
\usepackage{tikz}
\usepackage{multirow}
\usepackage[linesnumbered, ruled,vlined]{algorithm2e}
\usepackage{pdflscape}
% margin setup
\geometry{margin=0.8in}

@@

\allowdisplaybreaks

\title{Testing for the prevalence or transience of a shock effect in the post-shock setting}
\author{}
\date{}

@@

\maketitle
\begin{abstract}
 We provide a hypothesis test for the existence or transience of a shock in the post-shock setting of \cite{lin2021minimizing}.
\end{abstract}

\section{Setting}

@@

The assumed autoregressive models considered are that in \cite{lin2021minimizing}, we describe details here. Let $I(\cdot)$ be an indicator function, $T_i$ be the time length of the time series $i$ for $i = 1, \ldots, n+1$, and $T_i^*$ be the time point just before the one when the shock is known to occur, with $T_i^* < T_i$.  For $t= 1, \ldots, T_i$ and $i = 1, \ldots, n+1$, the model $\mc{M}_1$ is defined as
\begin{align}
\mc{M}_1 \colon y_{i,t} =\eta_i +\alpha_i D_{i,t} + \phi_i y_{i, t-1} + \theta_i'\mbf{x}_{i,t} + \varepsilon_{i,t}\label{equation1}
\end{align}
 where $D_{i,t} = I(t = T_i^* + 1)$ 
and $\x_{i,t} \in \R^{p}$ with $p \geq 1$.  We assume that the 
$\mbf{x}_{i,t}$s are fixed. Let $|x|$ denote the absolute value of $x$ for $x\in \reals$. For $i = 1, \ldots, n+1$ and $t=1, \ldots, T_i$, the random effects structure for $\mc{M}_1$ is:
\begin{align*}
  \eta_i &\simiid \mc{F}_{\eta} \text{ with }  \; \mrm{E}_{\mc{F}_{\eta}}(\eta_i) = 0, \mrm{Var}_{\mc{F}_{\eta}}(\eta_i)  = \sigma^2_{\eta}\\
  \phi_i &\simiid \mc{F}_{\phi} \text{ where } |\mc{F}_{\phi}| < 1, \\
   \theta_i &\simiid \mc{F}_{\theta} \text{ with }  \; \mrm{E}_{\mc{F}_{\theta}}(\theta_i) = \mu_{\theta}, \mrm{Var}_{\mc{F}_{\theta}}(\theta_i)  = \Sigma^2_{\theta} \\
\alpha_i &\simiid \mc{F}_{\alpha} \text{ with }  \; \mrm{E}_{\mc{F}_{\alpha}}(\alpha_i) = \mu_{\alpha}, \mrm{Var}_{\mc{F}_{\alpha}}(\alpha_i)  = \sigma^2_{\alpha}  \\
\varepsilon_{i,t} & \simiid  \mc{F}_{\varepsilon_i} \text{ with }  \; \mrm{E}_{\mc{F}_{\varepsilon_i}}(\varepsilon_{i,t}) = 0, \mrm{Var}_{\mc{F}_{\varepsilon_i}}(\varepsilon_{i,t})  = \sigma^2_i  ,  \\
\eta_i &\indep  \alpha_i \indep \phi_i \indep \theta_i \indep \varepsilon_{i,t},
\end{align*}
where $\indep $ denotes the independence operator. We also consider a modeling framework where the shock effects are linear functions of covariates with an additional additive mean-zero error. For $i = 1, \ldots, n+1$, the random effects structure for this model (model $\mc{M}_2$) is:
\begin{align}
\mc{M}_2 \colon \begin{array}{l}
  y_{i,t} =\eta_i +\alpha_i D_{i,t} + \phi_i y_{i, t-1} + \theta_i'\mbf{x}_{i,t} + \varepsilon_{i,t}\\[.2cm]
  \; \alpha_i = \mu_{\alpha}+\delta_{i}'\mbf{x}_{i, T_i^*+1}+ \t{\varepsilon}_{i},
\end{array}\label{model2}
\end{align}
 where the added random effects are
\begin{align*}
\t{\varepsilon}_{i} &\simiid  \mc{F}_{\t{\varepsilon}} \text{ with }\mrm{E}_{\mc{F}_{\t{\varepsilon}}}(\t{\varepsilon}_{i})=0, \mrm{Var}_{\mc{F}_{\t{\varepsilon}}}(\t{\varepsilon}_{i})=\sigma^2_{\alpha}\\
\eta_i &\indep  \alpha_i \indep \phi_i \indep \theta_i \indep \varepsilon_{i,t} \indep \t{\varepsilon}_{i}.
\end{align*} 
We further define 
$\tilde{\alpha}_i=\mu_{\alpha}+\delta_i'\mbf{x}_{i, T_i^*+1}$. 
We will investigate the post-shock aggregated estimators in $\mc{M}_2$ 
in settings where $\delta_i$ is either fixed or random. 
We let $\mc{M}_{21}$ denote model $\mc{M}_{2}$ with $\delta_i = \delta$ for $i= 1, \ldots, n+1$, 
where $\delta$ is a  fixed unknown parameter.
We let $\mc{M}_{22}$ denote model $\mc{M}_{2}$ with the following random effects :
\begin{align*}
&\delta_i\simiid  \mc{F}_{\delta} \text{ with }\mrm{E}_{\mc{F}_{\delta}}(\delta_i)=\mu_{\delta}, \mrm{Var}_{\mc{F}_{\delta}}(\delta_i)=\Sigma_\delta \\
   & \delta_i  \indep \t{\varepsilon}_{i}.
\end{align*}
We further define the parameter sets
\begin{align}
\begin{array}{l}
 \;\, \Theta = \{(\eta_i, \phi_i, \theta_i, \alpha_i, \mbf{x}_{i,t}, y_{i,t-1}, \delta_i)\colon    t= 1, \ldots, T_i, i = 2, \ldots, n +1\} \\
   \Theta_1 = \{(\eta_i, \phi_i, \theta_i, \alpha_i, \mbf{x}_{i,t}, y_{i,t-1}, \delta_i)\colon  t= 1, \ldots, T_i, i = 1\}
\end{array}\label{parameter} 
\end{align}
where $\Theta$ and $\Theta_1$ can adapt to $\mc{M}_1$ by dropping $\delta_i$. We assume this for notational simplicity.

\subsection{Forecast}
\label{forecast}

In our post-shock setting we consider the following candidate forecasts: 
\begin{align*}
  &\text{Forecast 1}: \hat y_{1,T_1^*+1}^1 = \hat\eta_1 
    + \hat\phi_1 y_{1,T_1^*} + \hat\theta_1'\x_{1,T_1^*+1} 
    , \\
  &\text{Forecast 2}: \hat y_{1,T_1^*+1}^2 = \hat\eta_1 
    + \hat\phi_1 y_{1,T_1^*} + \hat\theta_1'\x_{1,T_1^*+1} 
    + \hat{\alpha},
\end{align*}
where $\hat\eta_1$, $\hat\phi_1$, and $\hat\theta_1$ are all OLS estimators of $\eta_1$, $\phi_1$, and $\theta_1$, respectively, and $\hat{\alpha}$ is some form of estimator for the shock effect of time series of interest, i.e., $\alpha_1$. 
%The first forecast ignores the presence of $\alpha_1$ while the second forecast incorporates an estimate of $\alpha_1$ that is obtained from the other independent forecasts under study.

Throughout the rest of this article we highlight when the information from the time series donor pool, indexed by $\{y_{i,t} \colon t = 1,\ldots,T_i, i = 2,\ldots,n+1\}$, can be used to construct a shock effect estimator $\hat\alpha$ in which Forecast 2 beats Forecast 1 and vice-versa. We will consider the different dynamic panel models $\mc{M}_1$, $\mc{M}_{21}$, and $\mc{M}_{22}$.
%Improvement will be measured by assessing the reduction in risk that Forecast 2 offers over Forecast 1. We outline the theoretical details of risk-reduction in Section \ref{properties}. We specifically focus on predictions for $y_{1,T_1^*+1}$, the first post-shock response. It is important to note that, in general, $\hat{\alpha}$ does not converge to $\alpha_1$ in any sense.  Despite this shortcoming, adjustment of the forecast for $y_{1,T_1^*+1}$ through the addition of $\hat{\alpha}$ has the potential to lower forecast risk under several conditions, and for several estimation techniques.
We want to determine which forecast is appropriate over a horizon while the methods in \cite{lin2021minimizing} were only appropriate in the nowcasting setting in which prediction was only focused on the response immediately following the shock.

\subsection{Multi-horizon forecast evaluation}

We now consider an approach for comparing the forecasts over a specified time horizon after a shock has occurred by borrowing information across similar forecasts.

\cite{quaedvlieg2021multi} provided a methodology for comparing forecasts jointly across all horizons of a forecast path, $h = 1,\ldots, H$. In our post-shock setting, we want to compare the forecasts 
$$
  \hat y^{1,h}_{1,t} \; \text{and} \; \hat y^{2,h}_{1,t}
$$
where $y^{1,h}_{1,t}$ is the forecast for $y_t$ that accounts for the yet-to-be observed structural shock and is based on the information set $\mathcal{F}_{t-h}$, and $\hat y^{2,h}_{1,t}$ is defined similarly for the forecast that does not include any shock effect information. We will compare these forecasts in terms of their loss differential
$$
  d_t = L_{1,t} - L_{2,t},
$$
where $L_{j,t} = L(y_t,\hat y_{j,t})$, $j = 1,2$, and $L$ is a loss function. Hypothesis tests in \cite{quaedvlieg2021multi} are with respect to $E(d_t) = \mu_t$. Conditions for these tests require conditions of \cite{giacomini2006tests}.

\textbf{Note}: We need more formality for constructing $\hat y^{1,h}_{1,t}$. We could use the forecasts in \cite{lin2021minimizing} and then consider $h$-ahead methods after adjusting for the shock. Or we could consider aggregation approaches which average all post-shock responses of the series in the donor pool.

We will consider the average superior predictive ability (aSPA) to assess whether or not a shock is permanent or transitory. The aSPA investigates forecast comparisons based on their weighted average loss difference
$$
  \mu^{(AVG)} = \textbf{w}^T\mathbf{\mu} = \sum_{h=1}^H w_h \mu^h
$$
with weights $\textbf{w}$ that sum to one. Note that aSPA requires the user to take a stand on the relative importance of under-performance at one horizon against out-performance at another, and note that it is likely that $\mu^h > 0$ for $h$ closer to 1 since the user expects that a structural shock will occur and the structural shock is taken into account by forecast 1.

\bibliographystyle{plainnat}
\bibliography{../synthetic-prediction-notes}
